---
title: "Reproducible Research week2"
output: html_document
---
#### Loading in the data.
```{r}
data <- read.csv("activity.csv",as.is = TRUE)


```

#### Remove N/A values and sum 

```{r}
data1 <- data[!is.na(data$steps),]

data1steps<- with(data1, tapply(steps, as.factor(data1$date), sum, na.rm = T))
```
### What is mean total number of steps taken per day?
```{R,echo=FALSE}
hist(data1steps,main = "Total number of steps taken per day", xlab = "Total number of steps")

summary(data1steps)
```
### What is the average daily activity pattern?
```{r,echo=FALSE}
data1stepsmean<- with(data1, tapply(steps, interval, mean, na.rm = T))

plot(names(data1stepsmean),data1stepsmean,type="l", xlab="Interval", ylab="# of Steps",main="Average Number of Steps per Day by Interval")

```

````{r, echo=FALSE}


data1stepsmeanF<-data.frame(interval=names(data1stepsmean),meansteps=data1stepsmean)
maxint<-which.max(data1stepsmeanF$meansteps)
maxintnumber<-data1stepsmeanF[maxint,1]


```


### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

## `r maxintnumber`


```{r, echo=FALSE}
missing<-sum(is.na(data))
```

### The total number of missing values is `r missing`


```{r,echo=FALSE}
avgsteps <- aggregate(steps ~ interval, data = data, FUN = mean)

FIXNA <- numeric()
for (i in 1:nrow(data)) {
    obs <- data[i, ]
    if (is.na(obs$steps)) {
        steps <- subset(avgsteps, interval == obs$interval)$steps
    } else {
        steps <- obs$steps
    }
    FIXNA <- c(FIXNA, steps)
}


dataclean <- data
dataclean$steps <- FIXNA
```
### What is mean total number of steps taken per day after N/A's are replaced

```{R,echo=FALSE}
datacleansteps<- with(dataclean, tapply(steps, as.factor(dataclean$date), sum, na.rm = T))
hist(datacleansteps,main = "Total number of steps taken per day", xlab = "Total number of steps")
```
```{r}
summary(datacleansteps)
```
The values do differ when N/A's are populated but the impact is only on the lower step intervals 

### Activity Patern Weekday and Weekend

```{r,echo=FALSE}

dates<-dataclean$date
dates<-as.Date(as.character(dates),"%Y-%m-%d")
dataclean2<-dataclean
dataclean2$day<-weekdays(dates)
dataclean2$date<-as.Date(as.character(dates),"%Y-%m-%d")
dataclean2$wend <- as.factor(ifelse(weekdays(dataclean2$date) %in% c("Saturday","Sunday"), "Weekend", "Weekday"))
avgstepsweekend <- aggregate(steps ~ interval + wend, data = dataclean2, FUN = mean)

```
```{r,echo=FALSE}

library(ggplot2)

panel<-ggplot(avgstepsweekend,aes(x=interval,y=steps,color=wend))+ facet_grid(wend~.)+geom_line()+ 
labs(title="Average Number of Steps Taken vs 5-min Interval on Weekday/Weekend", y="Average Number of Steps", x="5-min Interval Times Series")

panel

```